---
title: "Analysis"
author: "Laura Bock Paulsen & Daniel Blumenkranz"
date: "2023-03-27"
output: html_document
---

```{r}
pacman::p_load(raster, sf, tidyverse, ggplot2)
```

# Moran's I
The code in this section is made with inspiration from the following tutorial: https://mgimond.github.io/Spatial/spatial-autocorrelation-in-r.html

## Load in needed data
```{r}
cult <- read_csv("data/votes_cultural_similarity.csv", show_col_types = FALSE)
geo <- read_csv("data/votes_geo_nb.csv", show_col_types = FALSE)
votes <- read_csv("data/votes.csv", show_col_types = FALSE)
name_map <- read_csv("data/shortname_mapping.csv", show_col_types = FALSE)
```

## Define functions
```{r}
average_nbh_points <- function(data, points_col) {
  data$avg_nb <- 0
  
  for (i in 1:nrow(data)) {
    neighbours <- data[i,]$countries_similar
    neighbours <- strsplit(neighbours, " ")[[1]]
    
    # the points the country itself got
    points <- data[i,][[points_col]]
    
    # points the neighbours got
    neighbourhood_votes <- data %>% filter(country_sn %in% neighbours)
    
    data[i,]$avg_nb <- mean(c(neighbourhood_votes[[points_col]], points), na.rm = TRUE)
  }
  
  return(data)
}


simulate <- function(data, points_col, n){
  # Create an empty vector
  I.r <- vector(length=n) 
  
  tmp_data <- data
  
  for (i in 1:n){
    # Randomly shuffle income values
    x <- sample(data[[points_col]], replace=FALSE)
    
    tmp_data[[points_col]] <- x
    
    # Compute new set of lagged values
    tmp_data <- average_nbh_points(tmp_data, points_col)
    x.lag <- tmp_data$avg_nb
    
    # Compute the regression slope and store its value
    M.r    <- lm(x.lag ~ x)
    I.r[i] <- coef(M.r)[2]
  }
  
  return(I.r)
}


pseudo_pval <- function(beta, simulated_betas, n){
  N_greater <- sum(beta > simulated_betas)
  p_value <- min(N_greater + 1, n + 1 - N_greater) / (n + 1)
  
  return(p_value)
}


plot_morans_i <- function(data, neibourhood, individual, title){
  ggplot(data, aes(x = individual, y = neibourhood)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "darkblue") + 
  ggtitle(title) +
  theme_light()
}

```

## Calculating Moran's I and pseudo p-value
### Jury - Geographical neighbours

```{r}
point_column <- "points_nb_jury"
average_column <- "jury_avg_nb"
data1 <- geo
n_sim <- 500

M <- lm(data1[[average_column]] ~ data1[[point_column]])
plot_morans_i(data1, data1[[average_column]], data1[[point_column]], "Jury, geographical neighbours")
simulated_correlations <- simulate(data1, points_col = point_column, n = n_sim)

# Plot the histogram of simulated Moran's I values
# then add our observed Moran's I value to the plot
hist(simulated_correlations, main=NULL, xlab="Moran's I", las=1)
abline(v=coef(M)[2], col="red")

p <- pseudo_pval(coef(M)[2], simulated_correlations, n_sim)
paste("Moran's I coef:", coef(M)[2])
paste("psuedo_pval:", p)
```

### Tele - Geographical neighbours
```{r}
point_column <- "points_nb_tele"
average_column <- "tele_avg_nb"
data1 <- geo
n_sim <- 500

M <- lm(data1[[average_column]] ~ data1[[point_column]])
plot_morans_i(data1, data1[[average_column]], data1[[point_column]], "Tele, geographical neighbours")
simulated_correlations <- simulate(data1, points_col = point_column, n = n_sim)

# Plot the histogram of simulated Moran's I values
# then add our observed Moran's I value to the plot
hist(simulated_correlations, main=NULL, xlab="Moran's I", las=1)
abline(v=coef(M)[2], col="red")

p <- pseudo_pval(coef(M)[2], simulated_correlations, n_sim)
paste("Moran's I coef:", coef(M)[2])
paste("psuedo_pval:", p)
```

### Jury - Cultural neighbours
```{r}
point_column <- "points_cult_similar_jury"
average_column <- "jury_avg_nb"
data1 <- cult
n_sim <- 500

M <- lm(data1[[average_column]] ~ data1[[point_column]])
plot_morans_i(data1, data1[[average_column]], data1[[point_column]], "Jury, cultural neighbours")

simulated_correlations <- simulate(data1, points_col = point_column, n = n_sim)

# Plot the histogram of simulated Moran's I values
# then add our observed Moran's I value to the plot
hist(simulated_correlations, main=NULL, xlab="Moran's I", las=1)
abline(v=coef(M)[2], col="red")

p <- pseudo_pval(coef(M)[2], simulated_correlations, n_sim)

paste("Moran's I coef:", coef(M)[2])
paste("psuedo_pval:", p)
```

### Tele - Cultural neighbours
```{r}
point_column <- "points_cult_similar_tele"
average_column <- "tele_avg_nb"
data1 <- cult
n_sim <- 500

M <- lm(data1[[average_column]] ~ data1[[point_column]])
plot_morans_i(data1, data1[[average_column]], data1[[point_column]], "Tele, cultural neighbours")
simulated_correlations <- simulate(data1, points_col = point_column, n = n_sim)

# Plot the histogram of simulated Moran's I values
# then add our observed Moran's I value to the plot
hist(simulated_correlations, main=NULL, xlab="Moran's I", las=1)
abline(v=coef(M)[2], col="red")

p <- pseudo_pval(coef(M)[2], simulated_correlations, n_sim)
paste("Moran's I coef:", coef(M)[2])
paste("psuedo_pval:", p)
```

# PLOTS
## Preparing country polygons
```{r}
library(tmap)
library(RColorBrewer)
```

```{r}
countries <- getData("countries")

countries <- st_transform(st_as_sf(countries), crs = 3035)

# rename some countries
countries$NAME_ENGLISH <- gsub("Macedonia", "North Macedonia", countries$NAME_ENGLISH)
countries$NAME_ENGLISH <- gsub("\\band\\b", "&", countries$NAME_ENGLISH)
```

```{r}
# define bbox and use to crop countries
bbox <- st_bbox(c(xmin = 2400000, xmax = 8000000, ymin = 1000000, ymax=6000000), crs = st_crs(3035))

countries_crop <- st_crop(countries, bbox)

# simplify
countries_crop_simp <- st_simplify(st_make_valid(countries_crop), dTolerance=5000, preserveTopology = TRUE)

# join geo and cult
countries_full <- countries_crop_simp %>% 
                  left_join(geo, by = c("NAME_ENGLISH"="country")) %>% 
                  left_join(cult, by = c("NAME_ENGLISH"="country")) 

# prepare dataframe for facetting
countries_full_facet <- rbind(countries_full, countries_full, countries_full, countries_full)

# add point type indicator column
countries_full_facet$plot <- c(rep("Cultural neighbours: Jury",78), rep("Cultural neighbours: Televoters",78), rep("Geographic neighbours: Jury", 78), rep("Geographic neighbours: Televoters", 78))

# combine points in same order
countries_full_facet$points <- c(countries_full$points_cult_similar_jury, countries_full$points_cult_similar_tele, countries_full$points_nb_tele, countries_full$points_nb_jury)
```


## tmap plot, TODO: Explain missing countries

```{r}
tm_shape(countries_full_facet) +
tm_polygons(col = "points",
            palette = '-Spectral',
            style = "cont",
            title = 'Neigbour bias score', 
            border.col = 'white') +
tm_layout(bg.color = "lightblue",
          legend.position = c('RIGHT', 'TOP'),
          legend.bg.color = "white",
          inner.margins = c(0,0,0,0),
          panel.labels = c('Cultural neighbours: Jury',
                           'Cultural neighbours: Televoters',
                           'Geographic neighbours: Jury',
                           'Geographic neighbours: Televoters'),
          fontfamily = 'sans',
          panel.label.bg.color = 'white',
          panel.label.size = 1.2) +
tm_facets("plot", ncol=2,
          inside.original.bbox = TRUE) +  
tm_compass(position = c('RIGHT', 'BOTTOM')) +
tm_scale_bar(position = c('LEFT', 'BOTTOM'))
```

## ggplot

```{r}
ggplot() +
  geom_sf(data = countries_full_facet, aes(fill=points), color='white', lwd=0.02) +
  facet_wrap(.~plot) +
  theme_bw() +
  theme(strip.background = element_rect(fill='white'),
        panel.background = element_rect(fill = 'lightblue'),
        panel.grid.major = element_line(size = 0.1),
        axis.text.x = element_text(size=7),
        axis.text.y = element_text(size=7)) +
  scale_fill_gradientn(colors = brewer.pal(n=10, name='Spectral')[7:1], na.value = '#bdbfbe', name = "Neighbour bias score") +
  coord_sf(expand = FALSE)

```

